---
title: "Prediction Assignment Project"
author: "Diego Mora"
date: "6/13/2019"
output: html_document
---

## Summary

In this report we want to construct a model for predict the correct exceution of a particular excercise using the data of some sensors applied to 6 athletes. We are going to load and transform the data and then fit a random forest model.

```{r, include=FALSE, echo=TRUE}
library(dplyr)
require(ggplot2)
require(caret)
require(randomForest)
```
## Load the Data

```{r, cache=TRUE}
urltraining <- c("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv")
urltesting <- c("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv")

download.file(urltraining,paste(getwd(),"/training.csv",sep = ""))
download.file(urltesting,paste(getwd(),"/testing.csv",sep = ""))

training <- read.csv("training.csv", na.strings = c("NA","","#DIV/0!"))
testing <- read.csv("testing.csv", na.strings = c("NA","","#DIV/0!"))
```

```{r}
plot(training$classe)
```

## Transform

Select just the variables that have relevant data, it means the variables with less than 70% of NA values.

```{r}
training <- training %>% select(which(colMeans(is.na(.)) < 0.7))
```

As we don't want that our model depends on a particular participant we can take off the name variable, the time stamps and the "x" variable that is nothing more than the index.

```{r}
training <- training %>% select(6:60)
```

## Prediction Model

Data Partition:
```{r}
set.seed(1919)
inTrain <- createDataPartition(y=training$classe, p=0.7, list=FALSE)
trainingMod <- training[inTrain,]
testingMod <- training[-inTrain,]
```

First step is preprocessing the data using PCA and setting the threshold to find how many components we need. We want threshold or 95%, to find how many number of components are needed to capture the 95% of the variance. Note: we exclude the "classe" variable in the preprocess.

```{r}
preProc <- preProcess(trainingMod[,1:54],method = "pca",thresh = 0.95)
preProc
```

Now we can use our preprocessing to set a new training data with our Principal Components

```{r}
trainPC <- predict(preProc, trainingMod[,1:54])
```

We can now fit a model, we want to use Random Forest

```{r, cache=TRUE}
modelFit <- randomForest(trainingMod$classe~.,data = trainPC)
```

Our PCA for the testing set and then the predictions. We can meassure our model with the confusion matrix.

```{r, cache=TRUE}
testPC <- predict(preProc,testingMod[,1:54])
pred <- predict(modelFit,testPC)
confusionMatrix(testingMod$classe,predict(modelFit,testPC))
```

## Conclusion

This random forest model has a 97.5% of Accuracy with a 95% confidence interval of 97% to 97.9%. we need 25 Principal Components to describe the 95% of the variance of the data.

## References

The data used to this project was taken from:

Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.
